Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference Imaging Photometric Survey.}

### 2021-07-02 cluster_paper/ms.tex
update ms

%\documentclass[12pt,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{892} % plot_autorotation.py, Set A, with manual Prot-color cut
\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
\newcommand{\ncompfrac}{38.0\%} % plot_compstar_rotation.py, based on plot_autorotation
\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we use three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available
\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
accuracy of the Gaia-derived results are the current state of the art.
We therefore adopted what we viewed as the most important Gaia-based
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of these samples is
available online.\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}
%While we considered performing our own clustering analysis on the
%Gaia data, such an effort would in effect only replicate the work of
%these investigators.  We opt to instead use their studies as a starting
%point.

@@

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}

@@

We then measured the rotation period from the resulting light curve using
the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
the Stellingwerf phase-dispersion minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{\citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light from
the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5
pixels.  We recorded the top five periodogram peaks\deleted{ from each
method}, and their corresponding powers.  Finally, as a check on
crowding, we recorded the number of stars within the aperture with
greater brightness than the target star, and with brightness within 1.25
and 2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
$\mathcal{A}$}, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
light curves, out of \nautorotdenominator\ stars for which rotation
periods might plausibly have been detected.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle \replaced{and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%}{period
fell below the boundary drawn in Figure~\ref{fig:rot}, which was
constructed the filter using the Pleiades polynomial from
\citet{curtis_rup147_2020}, with hard cuts at the red and blue ends.}.
This yielded \nautorotnumeratormatching\ light curves\added{in Set
$\mathcal{B}$}. 
Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
set of field stars was also analyzed for comparison.  The results are
discussed in Appendix~\ref{app:compstar}.
\added{The justification for the additional cut in Set $\mathcal{B}$ is
that \cn\ is already known to have few if any stars above the slow
sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
are most likely therefore either not cluster members, or they are cluster
members, but with unresolved binary companions contaminating the rotation
period measurement \citep[{\it
e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
assessing which are rotationally consistent with being in the cluster,
this is a viable filter; for the purpose of fully exploring whether any
rotational outliers could be bonafide cluster members, Set $\mathcal{A}$
would be the better cut.}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
    \includegraphics[width=0.700\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
  $\mathcal{B}$ is shown} as in
  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
  above an empirical isochrone; 20\% of stars) are overplotted in
  orange.  
  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
  the fraction of stars showing signs of binarity in the slow and fast
  subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
  \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
  \label{fig:binarity}
	}
\end{figure*}

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space\added{ (from Set~$\mathcal{A})$}.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our TESS
pipeline succeeded in making light curves.  In other words, the stars in
the base sample are those for which we could have plausibly detected a
rotation period.  This color range is preferred because the the slow
sequence becomes less defined for spectral types later than $\approx$K4V.
The bins in the histograms (top-right) were chosen to ensure that the
same number of stars were in each bin.  Rotation periods
\added{consistent with a gyrochronological age of 150\,Myr} are detected
for \replaced{66\%}{68\%} of stars \replaced{at least 25\,pc from the
cluster core}{reported to be in the cluster}.  The average detection rate
inside of 25\,pc, \replaced{73\%}{72\%}, is slightly higher\added{ than
outside of 25\,pc, where the detection rate is 63\%}. Within
$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\%} is
relatively low, due to crowding and saturation in the TESS images.

@@

Even with this level of contamination, our rotation period
measurements show that the halo extends to separations of
$\approx$250\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400-500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats (orange
circles) visible in the top-left panels of
Figure~\ref{fig:physical_x_rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.9\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.700cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all candidate \cn\ members with Gaia-ESO
  ($N=459$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Gray
  points are field stars with Kepler planets
  \citep{berger_identifying_2018}.  Pleiades stars (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in yellow
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections; all stars 
  later than $\approx$M0 do not have significant lithium.
  The Li measurements for both core and halo stars in \cn\ are
  consistent with a near-Pleiades age.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  candidate \cn\ members with TESS rotation periods and Gaia-ESO or
  GALAH spectra, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ are shown with gray
  crosses.  The lithium-rotation correlation for the K-dwarfs has been
  observed in other young clusters, and is discussed in
  Section~\ref{discussion:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

@@

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} demonstrated that between 100 and
200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
the Pleiades) all show similar rotation period distributions.  The FGK
stars show a ``slow sequence'', a population of ``rapid rotators'',
and in some cases stars in the ``void'' between the two populations
({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
Pleiades).  The late M dwarfs with detected rotation periods tend to
rotate rapidly ($<3$\,days), though there may also be late M dwarfs
with slow ($>20$\,days) rotation periods
\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
beyond the magnitude limit of our analysis, they have been explored in
depth by \citet{Irwin_NGC2516_2007} and
\citet{fritzewski_rotation_2020}.  A comparison of our results is
shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
includes, in order of precedence: rotation periods from our analysis,
from \replaced{``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
from the rotators in \citet{Irwin_NGC2516_2007}}{``Class 1 and
2'' \citet{fritzewski_rotation_2020} sources, from
Table 1 of \citet{healy_stellar_2020}, and
from \citet{Irwin_NGC2516_2007}.}.  The main
contribution of our study beyond these previous analyses is our
sensitivity across both the core and halo of \cn, and a corresponding
expansion in the number of stars that can be analyzed.  In the
following section, we therefore focus on the origins of the features
in the rotation-color diagram, and whether our expanded sample can
provide any new insights.

@@

\paragraph{Age from Gyrochronology}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrpo<1.2$.  At redder colors, $1.2<\bpmrpo<1.7$, the
dispersion in rotation periods increases.  The maximum rotation
periods seen in \cn\ across both Sets $\mathcal{A}$ and $\mathcal{B}$
extend up to $\approx$11 days, rather than the $\approx$8.5 day upper
limit seen in the Pleiades.  This is consistent with \cn\ being
slightly older than the Pleiades.

@@

    $\approx$50\%.  Given the number density ratios of cluster to
    field stars at the widest separations ($\approx 10^{-3}$,
    \citealt{meingast_2021}), these contamination fractions are not as
    bad as one might fear.

@@

We performed the same light curve stitching and period-search
procedure discussed in the text on the field stars.
Imposing the same requirements for crowding resulted in
\ncompstardenominator\ stars for which rotation periods could have
been detected.  Imposing the same Lomb Scargle power cutoff (${\rm
LSP}>0.08$) and period upper limit ($P<15\,{\rm days}$) yielded
\ncompstarnumerator\ period detections (\ncompfrac).  Within the same
brightness cutoff, \nautovscompstarnumerator\ of
\nautovscompstardenominator\ cluster stars yielded period detections
(\nautofrac).  While the detection fractions are only mildly
different (likely due to the brightness cutoff), the period--color
distributions are very different
(Figure~\ref{fig:compstar}).
This confirms our expectation that our period measurement procedure is
not biased in any way that could produce the features we observe in
\cn.

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{892} % plot_autorotation.py, Set A, with manual Prot-color cut

%\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
%\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py

\newcommand{\ncompstardenominator}{656} % plot_compstar_rotation.py BpmRp0 from 0.4 to 0.8
%\newcommand{\ncompstarnumerator}{223} % plot_compstar_rotation.py, ditto, with Prot detected in field star sample
\newcommand{\ncompstarnumerator}{107} % plot_compstar_rotation.py, ditto, with Prot meeting 'set B' criteria in field star sample
\newcommand{\ncompfrac}{16.3\%} % plot_compstar_rotation.py, based on plot_autorotation for defaultcleaning_cutProtColor

\newcommand{\nautovscompstardenominator}{327} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{164} % plot_autorotation.py
\newcommand{\nautofrac}{50.1\%} % plot_autorotation.py

@@

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we apply three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available
\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
accuracy of the Gaia-derived results are the current state of the art.
We therefore adopted what we viewed as the most important Gaia-based
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of these samples is
available online.\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}, last accessed \texttt{2021/07/01}.}
%While we considered performing our own clustering analysis on the
%Gaia data, such an effort would in effect only replicate the work of
%these investigators.  We opt to instead use their studies as a starting
%point.

@@

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotation periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.480\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
			$\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
			$\mathcal{B}$ is shown} as in
		Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
		11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
		above an empirical isochrone; 20\% of stars) are overplotted in
		orange.  
		Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
		Dividing these stars into ``slow'' and ``fast'' sequences by eye,
		the fraction of stars showing signs of binarity in the slow and fast
		subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
		\replaced{56\% (40/71)}{51\% (68/134)}, respectively.
		\label{fig:binarity}
	}
\end{figure}

@@

We then measured the rotation period from the resulting light curve using
the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
the Stellingwerf phase-dispersion minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{ \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light from
the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5
pixels.  We recorded the top five periodogram peaks\deleted{ from each
method}, and their corresponding powers.  Finally, as a check on
crowding, we recorded the number of stars within the aperture with
greater brightness than the target star, and with brightness within 1.25
and 2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
$\mathcal{A}$}, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
\replaced{light curves}{stars}, out of \nautorotdenominator\ stars \replaced{for which rotation
periods might plausibly have been detected}{that met the initial crowding requirements}.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle \replaced{and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%}{period
fell below the boundary drawn in Figure~\ref{fig:rot}, which was
constructed using the Pleiades polynomial from
\citet{curtis_rup147_2020}}.
This yielded \nautorotnumeratormatching\ light curves. 
Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
set of field stars was also analyzed for comparison.  The results are
discussed in Appendix~\ref{app:compstar}.
\added{The rationale behind the additional cut in Set $\mathcal{B}$ is
that \cn\ is already known to have few if any stars above the slow
sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
are therefore most likely either not cluster members, or they are cluster
members, but with unresolved binary companions contaminating the rotation
period measurement \citep[{\it
e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
assessing which stars are rotationally consistent with being in the cluster,
this is a viable filter; for the purpose of exploring whether any
rotational outliers might be bonafide cluster members, Set $\mathcal{A}$
would be the better cut.}

@@

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods \added{in Set $\mathcal{B}$} is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space\added{ (from Set~$\mathcal{A})$}.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our TESS
pipeline succeeded in making light curves.  In other words, the stars in
the base sample are those for which we would have expected to detect a
rotation period.  This color range is preferred because the the slow
sequence becomes less defined for spectral types later than $\approx$K4V.
The bins in the histograms (top-right) were chosen to ensure that the
same number of stars were in each bin.  Rotation periods
\added{consistent with a gyrochronological age of 150\,Myr} are detected
for \replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc from the
cluster core}{reported to be in the cluster}.  The average detection rate
inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is slightly higher\added{ than
outside of 25\,pc, where the detection rate is 63\% (181/288)}. Within
$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\% (26/44)} is
\replaced{relatively}{anomalously} low, due to crowding and saturation in the TESS images.

@@

\replaced{Even with}{Despite} this level of contamination, the rotation period
measurements show that the halo extends to separations of
$\approx$250\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400-500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats (orange
circles) visible in the top-left panels of
Figure~\ref{fig:physical_x_rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.95\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.700cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all candidate \cn\ members with Gaia-ESO
  ($N=459$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Gray
  points are field stars with Kepler planets
  \citep{berger_identifying_2018}.  Pleiades stars (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in yellow
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections; all stars 
  later than $\approx$M0 do not have significant lithium.
  The Li measurements for both core and halo stars in \cn\ are
  consistent with a near-Pleiades age.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
		candidate \cn\ members with TESS rotation periods\added{ (Set $\mathcal{A}$)} and Gaia-ESO or
		GALAH spectra, and are colored by equivalent width.  Points with
		reported equivalent widths below 20\,m\AA\ are shown with gray
		crosses.  The lithium-rotation correlation for the K-dwarfs has been
		observed in other young clusters, and is discussed in
		Section~\ref{discussion:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} demonstrated that between 100 and
200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
the Pleiades) all show similar rotation period distributions.  The FGK
stars show a ``slow sequence'', a population of ``rapid rotators'',
and in some cases stars in the ``void'' between the two populations
({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
Pleiades).  The late M dwarfs with detected rotation periods tend to
rotate rapidly ($<3$\,days), though there may also be late M dwarfs
with slow ($>20$\,days) rotation periods
\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
beyond the magnitude limit of our analysis, they have been explored in
depth by \citet{Irwin_NGC2516_2007} and
\citet{fritzewski_rotation_2020}.  A comparison of our results is
shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
includes, in order of precedence: rotation periods from our analysis,
from \replaced{``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
from the rotators in \citet{Irwin_NGC2516_2007}}{``Class 1 and
2'' \citet{fritzewski_rotation_2020} sources, from
Table 1 of \citet{healy_stellar_2020}, and
from \citet{Irwin_NGC2516_2007}}.  The main
contribution of our study beyond these previous analyses is our
sensitivity across both the core and halo of \cn, and a corresponding
expansion in the number of stars that can be analyzed.  In the
following section, we therefore focus on the origins of the features
in the rotation-color diagram, and whether our expanded sample can
provide any new insights.

@@

\paragraph{Age from Gyrochronology}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrpo<1.2$.  At redder colors, $1.2<\bpmrpo<1.7$, the
dispersion in rotation periods increases.  The maximum rotation
periods seen in \cn\ \deleted{across both Sets $\mathcal{A}$ and
$\mathcal{B}$}
extend up to $\approx$11 days, rather than the $\approx$8.5 day upper
limit seen in the Pleiades.  This is consistent with \cn\ being
slightly older than the Pleiades.

@@

    $\approx$50\%.  \replaced{Given the number density ratios of cluster to field stars at the widest separations ($\approx 10^{-3}$, \citealt{meingast_2021}), these contamination fractions are not as bad as one might fear.}{While the latter contamination fraction may seem high, a comparison sample of field stars (Appendix~\ref{app:compstar}) has a detection rate three times lower, and a  distribution in the period-color plane inconsistent with that of a population with a single age.}

@@

\explain{Numbers in this paragraph have been updated per the revised
definition of Set $\mathcal{B}$. }
